---
title: "Analysis of Serotonin Elisa data sets in R"
output: html_notebook
---

## Motivation
- Elisa is a biological experiment that mimics dose response curves especially when it comes to the standardization procedure
- drc are often characterized by increase response with respect to increased dose and vice versa.
- The most impressive thing is this data was given to me without the concentration values so I had to critically think to determine what it took to analyze elisa data.
- Also it took me a while to know the objective of elisa data but we aim to determine the concentration of proteins within a sample by using the standard curve data. In this case the protein measured is serotomin.

## Key Lessons to self
- This data set was not generated by me but by an old colleague in the lab in my undergraduate institution so I was thinking of normzlization when i saw the data sets
- Normalization is often if not exclusively on data from the same sample as normalization then applies
- Normalization on data from different samples is not sound practice

### Import elisa data into R

```{r}
# import elisa data
# std has just the std expected OD values
# sorted has all the expected OD values
# Our job is using the std ODs to estimate ODs for samples
elisastd <- read.csv(file = 'elisastd.csv', header = T)
elisasorted <- read.csv(file = 'elisasorted.csv', header = T)
head(elisastd,8)
```

### Isolation of standard samples

```{r}
# define attributes in the data (a1= concentration (ng/ml),
# a2 = concentration (ng/ml),od = OD or optical density)

# What should the standard curve look like 
# OD @ ~450 nM (y) vs Conc ng/ml (x)

#isolate standard values only
elisastdonly <- head(elisastd)
elisastdonly <- elisastdonly[-c(1)]
```

### Plotting the standard curve

```{r}
plot_ly(elisastdonly, x = ~conc, y = ~odavg, type = 'scatter', mode = 'line', name = 'standard curve')
```

```{r}
#plotting the log of concentration for diagnostics
plot(log(elisastdonly$conc), elisastdonly$a2, pch = 21, 
     bg = 'black', ylab = 'Optical density', xlab = 'log(concentration)')
```

### Insisting on a linear model for value prediction
- You can use log of the concentration to stick with a linear model (y=mx+c)
- If not then we need to use the 4pL complex method
- Also this data is interesting because it did not follow the linear model as seen from the plotly diagram

## What is a 4 parameter logistics regression model?
      
    A more complex model than the linear mode. Simplicity and complexity 
    are opposites so this is NOT necessarily hard but complex. This means 
    hard and easy are opposite.
    Appropriate for dose response curves (drc) or receptor-ligand curves or
    assays as biologist will say. From a data perspective we use values as 
    proxies for the assay.
    
    